---
title: "Income Sufficiency"
author: "Bernhard Piskernik"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(plotly)
```
# constants

income related constants

```{r}
adult <- 4529
child <- 200

unemployed_factor <- 0.7

alimony <- c(0, 0.15, 0.25, 0.30) # alimony rate for 0, 1, 2, and 3  children
```


expenditure constants

```{r}
a1c0 <- 3977
a2c0 <- 6605
a2c1 <- 7057
a2c2 <- 7635
a2c3 <- 7968
```

For single parent households there is no differentiating between number of  children. Let's use analoguous value from two parent households and scale accordingly  after disentangling the costs for the adult and the children.

```{r}
a1cx <- 5354.18
a2cx <- 7492.73

a1c1 <-a1c0 + (a2c1-a2c0)/(a2cx-a2c0)*(a1cx-a1c0)
a1c2 <-a1c0 + (a2c2-a2c0)/(a2cx-a2c0)*(a1cx-a1c0)
a1c3 <-a1c0 + (a2c3-a2c0)/(a2cx-a2c0)*(a1cx-a1c0)
```

# create tables for incomes and expenditures

create data frame with incomes

```{r}
df_income <-crossing(
    employment_level = c(1, 0.75, 0.5),
    cohabiting_partner = c('none', 'working','not working'),
    children = 0:3
  ) %>% 
  mutate(
    income = adult * employment_level + child * children + case_when(
      cohabiting_partner == 'not working' ~ adult * unemployed_factor,
      cohabiting_partner == 'working' ~ adult,
      cohabiting_partner == 'none' ~ adult * alimony[children+1]
    ),
    adults = ifelse(cohabiting_partner == 'none', 1, 2)
  ) 
```

create data frame with the expenditures

```{r}
df_exp <- tibble(
    adults = rep(1:2, each=4),
    children = rep(0:3, 2)
  ) %>%
  # get expenditure my name pattern
  rowwise() %>% 
  mutate(
    exp = get(paste0('a', adults, 'c', children))
  ) 

```

join them and use the ratio of income to expenditure of a single person working full-time as reference to scale all other combinations

```{r}
df_final <- df_income %>%
  full_join(df_exp) %>%
  mutate(
    ratio = income/exp,
    sufficiency = ratio/(adult/a1c0),
  ) %>%
  arrange(cohabiting_partner, desc(employment_level))
```


```{r}

p  <-  df_final %>%
  ggplot(aes(x=children, y=sufficiency, fill=sufficiency, text=paste0(
      'income sufficiency:\n', round(sufficiency,3)
      ))) +
  geom_col() +
  facet_grid(employment_level ~ cohabiting_partner, labeller = labeller(
    cohabiting_partner = function(string) {paste0('cohabiting partner:\n', string)},
    employment_level = function(numeric) {paste0('employment:\n', as.numeric(numeric)*100,'%')}
  )) +
  guides(fill = 'none') +
  theme_minimal() +
  theme(plot.margin = margin(0.3,1,0.8,0.8, "cm")) +
  scale_fill_continuous(type = "viridis")

pp <- ggplotly(p, tooltip = "text") 
pp
```

```{r}
api_create(pp, filename = "income_sufficiency")
```

